---
title: "10x scRNA-seq Seurat pipeline for EPUc nEnd +/- 4OHT"
author: "Madeleine Linneberg-Agerholm"
date: '`r Sys.Date()`'
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_format='all',
                        output_dir='../reports/')})
output:
  # To create PDF report, uncomment below
  #pdf_document:
  #  toc: yes
  html_document:
    number_sections: yes
    theme: yeti
    toc: yes
    toc_float: yes
    df_print: paged
    dev: png
---

```{r knitr, include = FALSE}
DOCNAME = knitr::current_input()
knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.path     = paste0("cache/", DOCNAME, "/"),
                      cache.comments = TRUE,
                      echo           = TRUE,
                      error          = FALSE,
                      fig.align      = "center",
                      fig.path       = paste0("../reports/figures/", DOCNAME, "/"),
                      fig.width      = 10,
                      fig.height     = 8,
                      message        = FALSE,
                      warning        = FALSE)
```

# Introduction

QC, normalization, scaling and dimensionality reduction for whole dataset including 2iLIF, nEnd and 3D nEnd (originally named endoid)

# Load libraries

```{r message=FALSE}
library(tidyverse)
library(Seurat)
library(patchwork)

random_seed <- 12345
```

```{r message=FALSE}
data <- readRDS("00_EPUc_seurat_merged.rds")
cc <- readRDS("mouse_cell_cycle_genes.rds")
```

# QC

## Before filtering
```{r}
data[["percent.mt"]] <- PercentageFeatureSet(data, pattern = "^mt-")
VlnPlot(data, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```
```{r}
FeatureScatter(data, feature1 = "nCount_RNA", feature2 = "percent.mt")
FeatureScatter(data, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
```

```{r}
print(paste0("Before filtering: ", dim(data)[2], " cells ",  dim(data)[1], " genes"))
```

```{r}
nbins <- 100
min_genes <- 2000
max_genes <- 6500
min_cells <- 4000
max_cells <- 30000

ggplot(data@meta.data, aes(x=nCount_RNA)) + 
  geom_histogram(bins = nbins) + 
  geom_vline(aes(xintercept=min_cells), linetype="dashed", color='red') + 
  geom_vline(aes(xintercept=max_cells), linetype="dashed", color='red')

ggplot(data@meta.data, aes(x=nFeature_RNA)) + 
  geom_histogram(bins = nbins) + 
  geom_vline(aes(xintercept=min_genes), linetype="dashed", color='red') +
  geom_vline(aes(xintercept=max_genes), linetype="dashed", color='red')

ggplot(data@meta.data, aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mt)) + 
  geom_point() + 
  scale_color_continuous(type = "viridis") + 
  geom_vline(aes(xintercept=min_cells), linetype="dashed", color='red') +
  geom_vline(aes(xintercept=max_cells), linetype="dashed", color='red') +
  geom_hline(aes(yintercept=min_genes), linetype="dashed", color='red') +
  geom_hline(aes(yintercept=max_genes), linetype="dashed", color='red')
```
# Filter

```{r}
normalized_data <- subset(data, subset = 
                    nFeature_RNA > min_genes & nFeature_RNA < max_genes & 
                    nCount_RNA > min_cells & nCount_RNA < max_cells & 
                    percent.mt < 25)
```

```{r QC_after_filtering}
VlnPlot(normalized_data, 
        features = c("nFeature_RNA", "nCount_RNA", "percent.mt"),
        ncol = 4)
FeatureScatter(normalized_data, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
```

```{r}
print(paste0("After filtering: ", dim(normalized_data)[2], " cells ",  dim(normalized_data)[1], " genes"))
```

#Normalize data

```{r}
normalized_data <- NormalizeData(normalized_data)
normalized_data <- FindVariableFeatures(normalized_data)
```

##Identify the 10 most highly variable genes
```{r}
top10 <- head(VariableFeatures(normalized_data), 10)
```

##Plot variable features with and without labels
```{r}
plot1 <- VariableFeaturePlot(normalized_data)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot2
```
#Scaling the data

```{r}
all.genes <- rownames(normalized_data)
normalized_data <- ScaleData(normalized_data, features = all.genes)
```

#Perform linear dimensional reduction

```{r}
normalized_data <- RunPCA(normalized_data, features = VariableFeatures(object = normalized_data))
```

```{r}
ElbowPlot(normalized_data, ndims = 30)
```

```{r}
normalized_data <- CellCycleScoring(normalized_data, s.features = cc$s.genes, g2m.features = cc$g2m.genes, set.ident = FALSE)
```

```{r}
DimPlot(normalized_data, reduction = "pca", group.by = "Stage")
DimPlot(normalized_data, reduction = "pca", group.by = "Phase")
```

```{r}
normalized_data <- FindNeighbors(normalized_data, dims = 1:20)
normalized_data <- FindClusters(normalized_data, random.seed = 12345, resolution = 0.2)
normalized_data <- RunUMAP(normalized_data, reduction = "pca", dims = 1:20, seed.use = 12345)
```

# Save session

```{r}
saveRDS(normalized_data, file = "normalized_data_EPUc.rds")
```

# Session info

```{r session-info, cache = FALSE}
devtools::session_info()
```
